Lesson 6
========================================================

### Welcome
Notes:

***

### Scatterplot Review

```{r Scatterplot Review}
library(ggplot2) # setup the usual way
data(diamonds)
head(diamonds)
summary(diamonds)

ggplot(aes(x=carat, y=price), data=diamonds) + 
  geom_point(alpha=.25) + 
  xlim(0, quantile(diamonds$carat, probs=.99)) +
  ylim(0, quantile(diamonds$price, probs=.99)) +
  stat_smooth(method="lm")
```

***

### Price and Carat Relationship
Response:
As carat weight increases in intervals (mostly) of .5 carats, price varies wildly. 
***

### Frances Gerety
Notes: Great depression lessened diamond sales significantly as a luxury item

#### A diamonds is 
Forever


***

### The Rise of Diamonds
Notes:
Marketing included British royal family, movie stars etc.
***

### ggpairs Function
Notes:

```{r ggpairs Function}
# install these if necessary
install.packages('GGally')
install.packages('scales')
install.packages('memisc')
install.packages('lattice')
install.packages('MASS')
install.packages('car')
install.packages('reshape')
install.packages('plyr')

# load the ggplot graphics package and the others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)

# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ] # only pull random 10k instances to calculate on
ggpairs(diamond_samp, params = c(shape = I('.'), outlier.shape = I('.')))

```

What are some things you notice in the ggpairs output?
Response:
Price doesn't seem to correlate much with anything but carat weight
***

### The Demand of Diamonds
Notes:
Demand varies based on income. Those at the smaller end of the market cannot afford much, probably more competition,
so the variance of the prices is less, as well as the price range. Conversely, at the higher end market, the 
competition would not be as stiff, leading to increased varience. 

```{r The Demand of Diamonds}

library(gridExtra)

plot1 <- ggplot(aes(price), data=diamonds) + 
  geom_histogram(binwidth=50) +
  ggtitle('Price')

plot2 <- ggplot(aes(price), data=diamonds) +
  geom_histogram(binwidth=.01) +
  scale_x_log10() +
  ggtitle('Price (log10)')

grid.arrange(plot1, plot2, nrow=1)
```

***

### Connecting Demand and Price Distributions
Notes:
The peaks do seem to indicate that there are two specific target market, one for lower income individuals, and the other for much higher. The number sold spikes high sharply for lower priced diamonds as they approach a number from the lower side, but gradually taper off on the other side, which is the same for higher priced diamonds as well. 
***

### Scatterplot Transformation

```{r Scatterplot Transformation}

```


### Create a new function to transform the carat variable

```{r cuberoot transformation}
cuberoot_trans = function() trans_new('cuberoot', transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)
```

#### Use the cuberoot_trans function
```{r Use cuberoot_trans}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Overplotting Revisited

```{r Sort and Head Tables}
head(sort(table(diamonds$carat), decreasing = T))
head(sort(table(diamonds$price), decreasing = T))
```


```{r Overplotting Revisited}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point(position="jitter", aes(size=.75, alpha=.5)) + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Other Qualitative Factors
Notes:
The price seems to correlate a lot with cut as a secondary variable.
***

### Price vs. Carat and Clarity

Alter the code below.
```{r Price vs. Carat and Clarity}
# install and load the RColorBrewer package
install.packages('RColorBrewer', dependencies = TRUE) 
install.packages('RColorBrewer')
library(RColorBrewer)

ggplot(aes(x = carat, y = price, color=clarity), data = diamonds) + 
  geom_point( alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

***

### Clarity and Price
Response:
The price seems to correlate a lot with cut as a secondary variable.

***

### Price vs. Carat and Cut

Alter the code below.
```{r Price vs. Carat and Cut}
ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Cut', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')
```

***

### Cut and Price
Response:
Clarity doesn't seem to be much of a factor at all.
***

### Price vs. Carat and Color

Alter the code below.
```{r Price vs. Carat and Color}
ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Color', reverse = FALSE,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Color')
```

***

### Color and Price
Response:
Yes, color seems to affect the price in the same way that clarity did, which makes me think they're related somehow.
***

### Linear Models in R
Notes: lm() is used to build linear models
It takes a function in the following way: lm( f(x) ~ x )

Response:
The relationship we've seen with price in carat is: log10(price) ~ x^(1/3)
***

### Building the Linear Model
Notes:

```{r Building the Linear Model}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

***

### Model Problems
Video Notes:

Research:
(Take some time to come up with 2-4 problems for the model)
(You should 10-20 min on this)

Response:

One problem might be the lack of sample size in different cuts of diamonds. Most diamonds in the price model were higher quality cut, so the model may not be as effective when evaluating the price of lower cut diamonds. 

We might also want to think about price change over time. Is this sample size taken over a course of years? Decades? If so, it might have to adjust for inflation, or whether the market stays relatively the same i.e. does not become flooded. 

***

### A Bigger, Better Data Set
Notes:

```{r A Bigger, Better Data Set}
install.packages('bitops')
install.packages('RCurl')

library('bitops')
library('RCurl')

diamondsurl = getBinaryURL("https://raw.github.com/solomonm/diamonds-data/master/BigDiamonds.Rda")
load(rawConnection(diamondsurl))

load('diamonds-data-master/BigDiamonds.Rda')
```

The code used to obtain the data is available here:
https://github.com/solomonm/diamonds-data

## Building a Model Using the Big Diamonds Data Set
Notes:

```{r Building a Model Using the Big Diamonds Data Set}
diamondsbig.sub <- subset(diamondsbig, cert == 'GIA' & price < 10000 & !is.na(price))
diamondsbig.samp <- diamondsbig.sub[sample(1:length(diamondsbig.sub$price), 10000), ] # sample 10k of dataset

m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamondsbig.sample) #linear model comparing ln of price with cuberoot of carat
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)

suppressMessages(library(lattice))
suppressMessages(library(MASS))
suppressMessages(library(memisc))
models <- mtable(m1, m2, m3, m4, m5) # setup overall linear model contain all variables

# ggplot(aes(x = carat, y = price, color=clarity), data = diamondsbig) + # full dataset
ggplot(aes(x = carat, y = price, color=clarity), data = diamondsbig.samp) + # datasubset and sampled 10k
  geom_point( alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')



```


***

## Predictions

Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601

```{r}
#Be sure you’ve loaded the library memisc and have m5 saved as an object in your workspace.
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)
```

Evaluate how well the model predicts the BlueNile diamond's price. Think about the fitted point estimate as well as the 95% CI.

***

## Final Thoughts
Notes:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

